|

### 2024-07-11 doc/paper_mine.qmd
Finalised paper.qmd

Assignment I aims to replicate an empirical table that involves calculating the residual income of US firms over a defined period and examining the relationship between residual income and the Price-to-Book (P/B) ratio. The replication process includes data preparation, cleaning, normalization, and the application of statistical methods to compute and interpret financial metrics using Python. By following a step-by-step approach, I provide insights into the trends of residual earnings across different P/B groups, identical to trends presented by @Penman_2013. This document presents the analysis performed in Assignment I, adapted to the open science workflow. Note that intermediate check steps have been omitted.

@@

# Replication Steps

@@

# Check if files exist
if not os.path.exists(STATIC_DATA_PATH):
    raise FileNotFoundError(f"{STATIC_DATA_PATH} does not exist.")
if not os.path.exists(PANEL_DATA_PATH):
    raise FileNotFoundError(f"{PANEL_DATA_PATH} does not exist.")

# Load the data from Worldscope files into pandas DataFrames
static_data = pd.read_csv(STATIC_DATA_PATH, delimiter='\t')  # use '\t' due to tab-separation in the static file
panel_data = pd.read_excel(PANEL_DATA_PATH)

# Normalize column names to avoid issues caused by variation in column naming conventions
static_data.columns = static_data.columns.str.strip().str.lower().str.replace(' ', '_')
panel_data.columns = panel_data.columns.str.strip().str.lower().str.replace(' ', '_')

# Merge data on 'isin' to include country information in panel_data
merged_data = panel_data.merge(static_data[['isin', 'country']], on='isin', how='left')

# Print the merged data to check. Now we have added a column on countries in panel data
print("Step 1: Merged data sample on 'isin' from static and panel data:\n", merged_data.head(11))
print()  # add this to get a blank line for better readability
# To understand how many rows there are in the merged data
num_rows1 = len(merged_data)
print("Number of rows in merged data sample on 'isin' from static and panel data:", num_rows1)
print()
print()
```

```{python}
#| label: step2
#| echo: true
#| output: true

# Filter the merged data for US firms and years 1996-2015
us_firms = merged_data[merged_data['country'] == 'UNITED STATES']  # create a new DataFrame that contains only US firms
filtered_data = us_firms[(us_firms['year_'] >= 1996) & (us_firms['year_'] <= 2015)].copy()  # further filter the us_firms DataFrame to keep the task's required year range

# Print the filtered data to check
print("Step 2: Filtered data sample on US firms and years 1996-2015:\n", filtered_data.head(22))
print()

# To understand how many rows were dropped out
num_rows2 = len(filtered_data)
print("Number of rows in filtered data sample on US firms and years 1996-2015:", num_rows2)
print()
print()
```
```{python}
#| label: step3
#| echo: true
#| output: true

# Identify Year 0 to Year 6 for each firm within the filtered data
filtered_data['year_0'] = filtered_data.groupby('isin')['year_'].transform('min')  # this column will hold the according “Year 0” for each firm
filtered_data['relative_year'] = filtered_data['year_'] - filtered_data['year_0']  # this column now indicates how many years have passed since the firm’s “Year 0”
filtered_data = filtered_data[filtered_data['relative_year'] <= 6]  # filter the data to include only the first seven years (Year 0 to Year 6) for each firm

# Print the filtered data to check if it's empty or has data
print("Step 3: Filtered data sample after identifying the Year 0 and relative years:\n", filtered_data.head(22))
print()

# To understand how many rows were dropped out
num_rows3 = len(filtered_data)
print("Number of rows in filtered data sample after identifying the Year 0 and relative years:", num_rows3)
print()
print()
```

@@

# Calculate P/B Ratio for each firm in Year 0

year_0_data = filtered_data[filtered_data['relative_year'] == 0].copy()  # create a copy of filtered_data DataFrame to contain only data for Year 0
print("Step 4: Year 0 data:\n", year_0_data)
print()

year_0_data['bve'] = year_0_data['bve'].replace([0, np.inf, -np.inf], np.nan)  # replace zero and very small BVE values with NaN
# To verify: display rows where 'bve' is NaN
nan_bve_rows = year_0_data[year_0_data['bve'].isna()]
print("Rows where 'bve' has been replaced with NaN:\n", nan_bve_rows)
num_nan_bve_rows = len(nan_bve_rows)
print("Number of rows where 'bve' has been replaced with NaN:", num_nan_bve_rows)
print()

@@

# Before removing NaN P/B ratios
num_rows_before = len(year_0_data)
# Remove rows with NaN P/B ratios
year_0_data = year_0_data.dropna(subset=['p/b'])
# After removing NaN P/B ratios
num_rows_after = len(year_0_data)

# Calculate the number of dropped observations
num_dropped = num_rows_before - num_rows_after
print("Number of observations dropped due to NaN P/B ratios:", num_dropped)  # checked: the number of dropped observations is equal to number of dropped BVE observations
print("Number of rows in Year 0 data after removing NaN P/B ratios:", num_rows_after)
print("Year 0 data after calculating P/B ratio and dropping NaN values:\n", year_0_data)  # the DataFrame now contains new column p/b ratio
print()
print()
```

```{python}
#| label: step5
#| echo: true
#| output: true

# Form P/B groups

# Sort firms based on their P/B ratios for Year 0 in descending order
num_groups = 20  # define the number of groups
year_0_data = year_0_data.sort_values('p/b', ascending=False).reset_index(drop=True)  # sort the year_0_data DataFrame in descending order based on the P/B ratios
print("Step 5: Year 0 data sorted in descending order based on the P/B ratios:\n", year_0_data)  # check the correct order
print()

# Assign the groups so that the highest P/B ratios are in group 1 and lowest in group 20
year_0_data['p/b_group'] = pd.qcut(year_0_data.index, num_groups, labels=range(1, num_groups + 1))
print("Year 0 data updated with P/B groups:\n", year_0_data)  # checked: p/b ratios grouping aligns with previous output

# Count the number of firms in each P/B group before merging
pb_group_counts_pre_merge = year_0_data.groupby('p/b_group', observed=True)['isin'].nunique()
print("Number of firms assigned to each P/B group before merging:\n", pb_group_counts_pre_merge)
print()

# Calculate median P/B value for each group
median_pb_values = year_0_data.groupby('p/b_group', observed=True)['p/b'].median()
# Display the median P/B values
print("Median P/B values for each group:\n", median_pb_values)
print()

# Merge the P/B groups back to the filtered data for Year 0
filtered_data = filtered_data.merge(year_0_data[['isin', 'p/b_group']], on='isin', how='left')
# Fill missing P/B groups for years other than Year 0 with the same group as Year 0
filtered_data['p/b_group'] = filtered_data.groupby('isin')['p/b_group'].ffill()

# Display the combined filtered data after merging P/B groups
print("Filtered data after merging P/B groups:\n", filtered_data.head(22))  # the DataFrame now contains new column p/b_group
# Count the number of firms in each P/B group after merging
pb_group_counts_post_merge = filtered_data.groupby('p/b_group', observed=True)['isin'].nunique()
print("Number of firms assigned to each P/B group after merging:\n", pb_group_counts_post_merge)  # checked: same number as pre merging
# Calculate the number of rows in the merged filtered data sample
num_rows_merged_filtered = len(filtered_data)
print("Number of rows in the merged filtered data sample:", num_rows_merged_filtered)  # checked: same number as pre merging
print()
```

@@

# Calculate residual income for each P/B group for Years 0 to 6, deflated by the book value of equity at the end of the year before Year 0 (= beginning of year 0)

@@

# Create a copy of the merged data from Step 1 to preserve all years of observations
us_full_data = merged_data[merged_data['country'] == 'UNITED STATES'].copy()
print("Step 6: Merged data (Step 1) filtered to US only, all years preserved:\n", us_full_data.head(22))
print()

# Identify Year 0 for each firm, ensuring Year 0 is between 1996 and 2015
us_full_data['year_0'] = us_full_data.groupby('isin')['year_'].transform(lambda x: x[(x >= 1996) & (x <= 2015)].min())
us_full_data = us_full_data[us_full_data['year_0'].notna()]  # remove rows where year_0 is NaN
# Calculate the relative year and year_minus_1
us_full_data['relative_year'] = us_full_data['year_'] - us_full_data['year_0']
us_full_data['year_minus_1'] = us_full_data['year_0'] - 1

# Convert years to integer
us_full_data['year_0'] = us_full_data['year_0'].astype(int)
us_full_data['relative_year'] = us_full_data['relative_year'].astype(int)
us_full_data['year_minus_1'] = us_full_data['year_minus_1'].astype(int)
# Print the data to verify the steps
print("Data after identifying Year 0, relative year, and Year -1:\n", us_full_data.head(32))
print()

# Extract BVE for the year before Year 0 (relative_year = -1) for deflation
bve_deflation = us_full_data[us_full_data['relative_year'] == -1][['isin', 'bve']].rename(columns={'bve': 'bve_deflation'})
print(bve_deflation)
# Merge the bve_deflation values back to the filtered data
filtered_data = filtered_data.merge(bve_deflation, on='isin', how='left')
print("Filtered data with bve_deflation:\n",filtered_data)  # new column on bve_deflation
print()

# Calculate the BVE of the previous year for each year (for residual income calculation)
us_full_data['bve_prev'] = us_full_data.groupby('isin')['bve'].shift(1)
# Merge bve_prev back to filtered_data
filtered_data = filtered_data.merge(us_full_data[['isin', 'year_', 'bve_prev']], on=['isin', 'year_'], how='left')
print("Filtered data with bve_prev:\n",filtered_data)  # new column on bve_prev
print()

# Function to calculate non-deflated residual income
def calculate_residual_income(row):
    return row['ninc'] - (cost_of_equity * row['bve_prev'])
# Apply the function to calculate non-deflated residual income for each row
filtered_data['residual_income'] = filtered_data.apply(calculate_residual_income, axis=1) 
# Display the 'residual_income' values
print("Residual income values:\n", filtered_data)  # new column on residual_income
print()

# Function to deflate residual income
def deflate_residual_income(row):
    if pd.notna(row['bve_deflation']) and row['bve_deflation'] != 0:
        return row['residual_income'] / row['bve_deflation']
    else:
        return np.nan
# Apply the function to deflate residual income for each row
filtered_data['deflated_residual_income'] = filtered_data.apply(deflate_residual_income, axis=1)
print("Deflated residual income values:\n", filtered_data)  # new column on deflated_residual_income
print()
```

```{python}
#| label: step7
#| echo: true
#| output: true

# Output the replicated table

# Group by P/B group and relative year, then calculate the median deflated residual income for each group and year
median_residual_income = filtered_data.groupby(['p/b_group', 'relative_year'], observed=True)['deflated_residual_income'].median().unstack()

# Formatting the results (decimal places)
median_residual_income = median_residual_income.apply(lambda col: col.map(lambda x: f"{x:.3f}" if pd.notna(x) else "NaN"))
median_pb_values = median_pb_values.apply(lambda x: f"{x:.2f}")

@@

# Display the median residual income table
print("Median Residual Income by P/B Group and Relative Year:\n", median_residual_income)
print()
```

```{python}
#| label: step8
#| echo: true
#| output: true

# Step 8: Modification - exclude negative and very high P/B values

# Filter out extreme P/B values that are negative and higher than 7
filtered_year_0_data = year_0_data[(year_0_data['p/b'] > 0) & (year_0_data['p/b'] <= 7)]
print("Filtered Year 0 data after excluding negative and very high P/B values:\n", filtered_year_0_data)
print()

# Recalculate the P/B groups based on the filtered data
filtered_year_0_data = filtered_year_0_data.sort_values('p/b', ascending=False).reset_index(drop=True)
filtered_year_0_data['p/b_group'] = pd.qcut(filtered_year_0_data.index, num_groups, labels=range(1, num_groups + 1))
print("Filtered Year 0 data updated with P/B groups:\n", filtered_year_0_data)
print()

# Merge the P/B groups back to the filtered data for Year 0
filtered_data = filtered_data.drop(columns=['p/b_group'])
filtered_data = filtered_data.merge(filtered_year_0_data[['isin', 'p/b_group']], on='isin', how='left')
filtered_data['p/b_group'] = filtered_data.groupby('isin')['p/b_group'].ffill()

# Calculate median P/B values again
median_pb_values = filtered_year_0_data.groupby('p/b_group', observed=True)['p/b'].median()

# Group by P/B group and relative year, then calculate the median deflated residual income for each group and year
median_residual_income_filtered = filtered_data.groupby(['p/b_group', 'relative_year'], observed=True)['deflated_residual_income'].median().unstack()

# Formatting the results (decimal places)
median_residual_income_filtered = median_residual_income_filtered.apply(lambda col: col.map(lambda x: f"{x:.3f}" if pd.notna(x) else "NaN"))
median_pb_values = median_pb_values.apply(lambda x: f"{x:.2f}")

@@

# Display the median residual income table
print("Median Residual Income by P/B Group and Relative Year:\n", median_residual_income_filtered)
```

>>>>

Assignment I aims to replicate an empirical table that involves calculating the residual income of US firms over a defined period and examining the relationship between residual income and the Price-to-Book (P/B) ratio. The replication process includes data preparation, cleaning, normalization, and the application of statistical methods to compute and interpret financial metrics using Python. By following a step-by-step approach, I provide insights into the trends of residual earnings across different P/B groups, identical to trends presented by @Penman_2013. This document presents the analysis performed in Assignment I, adapted to the open science workflow. Note that intermediate check steps have been omitted for brevity. The full code, including intermediate print steps to verify outputs, is presented in `code/python/Assignment1_analyze_residual_income.py`.

@@

# Replication Steps
## Step 1: Data Loading and Merging
Begin by loading the data from the provided Worldscope files. The static data contains information on unique ISIN IDs for each firm, as well as the corresponding firm name and country. The panel data provides the financial performance numbers of each ISIN. Merge the data on the ‘isin’ column to include country information in the panel data. This merging step is crucial for filtering the US firms specifically in further steps.

@@

# Load the data from Worldscope files into pandas DataFrames
static_data = pd.read_csv(STATIC_DATA_PATH, delimiter='\t')
panel_data = pd.read_excel(PANEL_DATA_PATH)

# Normalize column names to avoid issues with column naming conventions
static_data.columns = (
    static_data.columns.str.strip()
    .str.lower()
    .str.replace(' ', '_')
)

panel_data.columns = (
    panel_data.columns.str.strip()
    .str.lower()
    .str.replace(' ', '_')
)

# Merge data on 'isin' to include country information in panel_data
merged_data = panel_data.merge(
    static_data[['isin', 'country']], 
    on='isin', 
    how='left'
)
```

## Step 2: Filter Data for US Firms and Relevant Years
Filter the merged data to include only US firms (excluding UK and German firms) and the year range 1996-2015. This ensures that further calculations are based on the specific subset of data that matches the assignment’s criteria.
```{python}
#| label: step2
#| echo: true
#| output: true

us_firms = merged_data[merged_data['country'] == 'UNITED STATES']  
filtered_data = us_firms[
    (us_firms['year_'] >= 1996) & 
    (us_firms['year_'] <= 2015)
].copy() 
```

## Step 3: Identify Year 0 to Year 6 for Each Firm
To calculate the P/B values and residual incomes for each firm/year, it is necessary to identify the base year (Year 0) and the subsequent relative years within the range 0 – 6 in the filtered data, as presented by Penman (2013). Define Year 0 as the earliest year in the dataset for each firm and display it in a new column ‘year_0’ that holds this value, making ‘year_0’ the same for each ISIN. Additionally, create the column ‘relative_year’, represented as the difference between the current year and Year 0. Since a firm may have more than 7 years of observations in the filtered data, limit the data to include only the first seven years (Year 0 to Year 6) for each ISIN to match the assignment’s requirements.
Notably, the number of observations after merging Step 1 (298,120), filtering Step 2 (142,837), and identifying years in Step 3 (89,584) nearly halves with each step.

```{python}
#| label: step3
#| echo: true
#| output: true

filtered_data['year_0'] = filtered_data.groupby('isin')['year_'].transform('min')  
filtered_data['relative_year'] = filtered_data['year_'] - filtered_data['year_0']  
filtered_data = filtered_data[filtered_data['relative_year'] <= 6]  
```

## Step 4: Calculate P/B Ratio for Each Firm in Year 0
According to Penman (2013), the residual earnings are calculated after P/B groups are formed in Year 0. Therefore, consider the firms’ financial performance data only for Year 0 to calculate the P/B ratio. Replace zero and infinite BVE values with NaN to drop invalid calculations
(affecting 28 observations in total). Use the following formula to calculate the P/B ratio:
$$
P/B_t = \frac{\textit{Market value of equity}_t}{\textit{Book value of equity}_t}
$$

@@

year_0_data = filtered_data[filtered_data['relative_year'] == 0].copy()

# replace zero and very small BVE values with NaN
year_0_data['bve'] = year_0_data['bve'].replace([0, np.inf, -np.inf], np.nan)

@@

# Remove rows with NaN P/B ratios
year_0_data = year_0_data.dropna(subset=['p/b'])
```

## Step 5: Form P/B Groups
Sort the firms based on their P/B ratios in descending order and divide them into 20 P/B groups, with the highest P/B ratios in group 1 and the lowest in group 20, as presented by Penman (2013). As a result, each P/B group contains approximately 900 firms, ensuring an equal division of the data set. Calculate the median P/B value for each group and merge the P/B group assignment information back into the filtered data, providing information on which P/B group each firm belongs to in each year.
```{python}
#| label: step5
#| echo: true
#| output: true

# Sort firms based on their P/B ratios for Year 0 in descending order
num_groups = 20  # define the number of groups
year_0_data = year_0_data.sort_values(
    'p/b', ascending=False
).reset_index(drop=True)

# Assign the groups so that the highest P/B ratios are in group 1
year_0_data['p/b_group'] = pd.qcut(
    year_0_data.index,
    num_groups,
    labels=range(1, num_groups + 1)
)

# Calculate median P/B value for each group
median_pb_values = year_0_data.groupby('p/b_group', observed=True)['p/b'].median()

# Merge the P/B groups back to the filtered data for Year 0
filtered_data = filtered_data.merge(
    year_0_data[['isin', 'p/b_group']],
    on='isin',
    how='left'
)

# Fill missing P/B groups for years other than Year 0 with same group as Year 0
filtered_data['p/b_group'] = filtered_data.groupby('isin')['p/b_group'].ffill()
```

## Step 6: Calculate Residual Income
This step involves calculating the residual income for each P/B group for Years 0 to 6, followed by deflation by the BVE at the beginning of Year 0. Assume a cost of equity capital of 8%, a reasonable assumption for US firms in the given time range. Based on Penman (2013), the residual income for each company/year should be deflated by the book value of equity at the beginning of Year 0.

To identify whether the financial data from Worldscope contains the numbers as of the beginning or end of the fiscal year, a random sample check was conducted. For example, for ISIN AN8068571086 in its *2020 Annual Report* (2021), the values for total assets, stockholders’ equity, net income, and other financial metrics are measured at the end of each fiscal year [@Schlumberger_2021, p. 20]. Hence, assume that the entire Worldscope panel data provides the BVE, MVE, and net income as values for the end of the respective fiscal year.

In case a company has observations before 1996, the filtered data set does not enable calculation of residual income in Year 0 (1996) due to missing value of BVE in Year -1, resulting into NaN. To avoid that, create a copy of the merged data from Step 1 to preserve all years of observations. Filter it to US companies and identify the year_minus_1 for each firm. Extract BVE for the year before Year 0 (relative_year = -1) for the deflation step and extract the BVE of the previous year for each year for the residual income calculation step.

Calculate the non-deflated residual income for each firm/year using the following formula:
$$
\text{Residual income}_t = \text{Net income}_t - r_e \times \text{Book value of equity}_{t-1}
$$

Assume that the beginning of Year 0 is equivalent to the end of the previous year because the financial data recorded at the end of one year represents the starting values for the following year. According to Penman (2013), residual income is deflated by book value at the beginning of year 0. Hence, for deflating the residual income, use the book value of equity from the year before Year 0 (bve_deflation) for all years (Years 1 to 6). This means that the BVE used for deflation remains constant across all years, normalizing the residual income across different firms and time periods. For example, for Year 0 (1996), use the BVE at the end of 1995 to deflate the residual income.

Calculate the deflated residual income for each firm/year using the following general formula:

$$
\text{Deflated residual income}_t = \frac{\text{Residual income}_t}{\text{Book value of equity}_{\text{Year 0 start}}}
$$

@@

@@

# Create a copy of the merged data from Step 1 to keep all years of observations
us_full_data = merged_data[merged_data['country'] == 'UNITED STATES'].copy()

# Identify Year 0 for each firm, ensuring Year 0 is between 1996 and 2015
us_full_data['year_0'] = us_full_data.groupby('isin')['year_'].transform(
    lambda x: x[(x >= 1996) & (x <= 2015)].min()
)
us_full_data = us_full_data[us_full_data['year_0'].notna()] 
# Calculate the relative year and year_minus_1
us_full_data['relative_year'] = us_full_data['year_'] - us_full_data['year_0']
us_full_data['year_minus_1'] = us_full_data['year_0'] - 1

# Convert years to integer
us_full_data['year_0'] = us_full_data['year_0'].astype(int)
us_full_data['relative_year'] = us_full_data['relative_year'].astype(int)
us_full_data['year_minus_1'] = us_full_data['year_minus_1'].astype(int)

# Extract BVE for the year before Year 0 (relative_year = -1) for deflation
bve_deflation = (
    us_full_data[us_full_data['relative_year'] == -1][['isin', 'bve']]
    .rename(columns={'bve': 'bve_deflation'})
)

# Merge the bve_deflation values back to the filtered data
filtered_data = filtered_data.merge(bve_deflation, on='isin', how='left')

# Calculate the BVE of the previous year for each year
us_full_data['bve_prev'] = us_full_data.groupby('isin')['bve'].shift(1)
# Merge bve_prev back to filtered_data
filtered_data = filtered_data.merge(
    us_full_data[['isin', 'year_', 'bve_prev']],
    on=['isin', 'year_'],
    how='left'
)

# Function to calculate non-deflated residual income
def calculate_residual_income(row):
    return row['ninc'] - (cost_of_equity * row['bve_prev'])
# Apply the function to calculate non-deflated residual income for each row
filtered_data['residual_income'] = filtered_data.apply(
    calculate_residual_income, axis=1
)

# Function to deflate residual income
def deflate_residual_income(row):
    if pd.notna(row['bve_deflation']) and row['bve_deflation'] != 0:
        return row['residual_income'] / row['bve_deflation']
    else:
        return np.nan
# Apply the function to deflate residual income for each row
filtered_data['deflated_residual_income'] = filtered_data.apply(
    deflate_residual_income, axis=1
)
```

## Step 7: Output the Replicated Table
Calculate the median deflated residual income for each P/B group and relative year. Include the median P/B values in the output table. Ensure consistency with Penman (2013) by formatting the results to three decimal places for residual income and two decimal places for P/B values. The resulting table will be saved into an Excel file in the `data/generated` folder of the repository.

The output table certainly indicates a wide range of P/B ratios among the firms, with some extreme values in high and low P/B groups. There is a noticeable trend where firms with very high or very low P/B ratios have less stable and more extreme residual earnings over time, while middle P/B groups seem to show more stable trends as presented by Penman (2013), suggesting a balanced reflection of expectations between market valuation and expected financial performance.

The variation in trends can be explained by the difference in data sets and possible additional assumptions made by Penman (2013) that are not available. For example, the assignment data set includes all USA firms from the Worldscope database, providing a broader sample than Penman’s original study, which included only NYSE and AMEX firms. Furthermore, the assignment analysis focuses on the years 1996 to 2015, a more recent period with different economic conditions compared to the time frame used in Penman’s (2013) study.
```{python}
#| label: step7
#| echo: true
#| output: true

median_residual_income = filtered_data.groupby(
    ['p/b_group', 'relative_year'], observed=True
)['deflated_residual_income'].median().unstack()

# Formatting the results (decimal places)
median_residual_income = median_residual_income.apply(
    lambda col: col.map(
        lambda x: f"{x:.3f}" if pd.notna(x) else "NaN"
    )
)
median_pb_values = median_pb_values.apply(lambda x: f"{x:.2f}")

@@

# Display the median residual income table
print(
    "Median Residual Income by P/B Group and Relative Year:\n", 
    median_residual_income
)
```

## Step 8: Modification - Exclude Extreme P/B Values
As indicated in Step 7, the output table certainly includes outliers. To replicate the more stable P/B values observed by Penman (2013), exclude extreme P/B values that are negative or higher than 7. Recalculate the P/B groups based on this filtered data. Merge the updated P/B group information back into the filtered data and repeat the calculation of median deflated residual income for each P/B group and relative year. Format the updated results with the required decimal places and save to an Excel file. The modified output of the replicated table, closely resembling Penman (2013), is presented in Table 2.

The modified table resembles more the one of Penman (2013) by presenting similar trends. In particular, the exclusion of extreme P/B values results in more stable and consistent residual earnings across P/B groups and years. High P/B groups show higher residual earnings initially, which slowly decline over time, and low P/B groups have negative residual earnings that tend to fluctuate but still remain negative.
```{python}
#| label: step8
#| echo: true
#| output: true

# Filter out extreme P/B values that are negative and higher than 7
filtered_year_0_data = year_0_data[
    (year_0_data['p/b'] > 0) & (year_0_data['p/b'] <= 7)
]

# Recalculate the P/B groups based on the filtered data
filtered_year_0_data = filtered_year_0_data.sort_values(
    'p/b', ascending=False
).reset_index(drop=True)

filtered_year_0_data['p/b_group'] = pd.qcut(
    filtered_year_0_data.index, num_groups, labels=range(1, num_groups + 1)
)

# Merge the P/B groups back to the filtered data for Year 0
filtered_data = filtered_data.drop(columns=['p/b_group'])
filtered_data = filtered_data.merge(
    filtered_year_0_data[['isin', 'p/b_group']], 
    on='isin', 
    how='left'
)
filtered_data['p/b_group'] = filtered_data.groupby('isin')['p/b_group'].ffill()

# Calculate median P/B values again
median_pb_values = filtered_year_0_data.groupby(
    'p/b_group', 
    observed=True
)['p/b'].median()

# Calculate the median deflated residual income for each group and year
median_residual_income_filtered = filtered_data.groupby(
    ['p/b_group', 'relative_year'], 
    observed=True
)['deflated_residual_income'].median().unstack()

# Formatting the results (decimal places)
median_residual_income_filtered = median_residual_income_filtered.apply(
    lambda col: col.map(lambda x: f"{x:.3f}" if pd.notna(x) else "NaN")
)
median_pb_values = median_pb_values.apply(lambda x: f"{x:.2f}")

@@

# Display the median residual income table
print(
    "Median Residual Income by P/B Group and Relative Year:\n",
    median_residual_income_filtered
)
```
